A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-09 shanghai.tex
Fixing proof of Corollary 10.3

We close the introduction with a more detailed description of the structure of this paper.
In Section~\ref{sec:category} we define quasicharacter sheaves and give an interpretation of
them as Weil sheaves with certain properties.
Section~\ref{sec:Frob} introduces the trace of Frobenius homomorphism,
one direction of the sheaf--function dictionary.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on the trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded and finite quasicharacter sheaves,
two subcategories of $\QC(G)$.  These alternate descriptions of quasicharacter sheaves play a key
role in the proof of Proposition~\ref{prop:restriction}.
In section \ref{sec:connected} we show that bounded and finite quasicharacter sheaves are the same
as standard quasicharacter sheaves when $G$ is connected, and that trace of Frobenius is an isomorphism.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and give a more concrete interpretation
of quasicharacter sheaves in this case.  Using a notion of global sections for quasicharacter sheaves,
we give isomorphisms between $\QCiso{G}$, $\Hh^1(\Weil{}, \Hom(G(\bFq), \EEx))$ and $\Hom(G(\Fq), \EEx)$.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
Section~\ref{sec:snake} closes out Part \ref{part1} with the proof of the main theorem: trace of Frobenius
is an isomorphism for $G$ as long as the geometric component group is finitely generated.

@@

Let $\Frob{}$ denote the geometric Frobenius element in $\Gal(\bFq/\Fq)$ as
well as the corresponding automorphism of $\Spec{\bFq}$. The Weil group
$\Weil{}\subset \Gal(\bFq/\Fq)$ is the subgroup generated by $\Frob{}$.
Let $\Frob{G} \ceq \id_{G} \times \Frob{}$ be the Frobenius automorphism of $\bG$.
When no other fields play a role we will write $\Frob{}$ and $\Weil{}$ instead.

@@

\begin{definition}\label{def:QC}
A \emph{quasicharacter sheaf on $G$} is a triple
$\qcs{L}\ceq (\gqcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(QC.1)}{QC.1} $\gqcs{L}$ is a rank~one $\ell$-adic local system on $\bG$, by which we mean
a constructible $\ell$-adic sheaf on $\bG$, locally constant on each connected component of $\bG$, whose stalks are one-dimensional $\EE$-vector spaces; 
\labitem{(QC.2)}{QC.2} $\mu: \bm^* \gqcs{L} \to \LxL$ is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gqcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gqcs{L} \boxtimes \gqcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gqcs{L} \boxtimes \bm^* \gqcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gqcs{L} \tight{0}{\boxtimes}{0} \LxL
  \end{tikzcd}
  \]
%
\labitem{(QC.3)}{QC.3} $\phi : \Frob{G}^* \gqcs{L} \to \gqcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gqcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gqcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gqcs{L}\boxtimes \Frob{G}^*\gqcs{L} \dar{\phi\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gqcs{L} \arrow{rr}{\mu}
    && \LxL
  \end{tikzcd}
  \]
\end{enumerate}
\end{definition}

Morphisms in the category of quasicharacter sheaves on $G$, denoted by $\QC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(QC.4)}{QC.4} if $\qcs{L} = (\gqcs{L},\mu,\phi)$ and
  $\qcs{L'} = (\gqcs{L'},\mu',\phi')$ are quasicharacter sheaves on $G$ then
  a morphism $\qcs{L} \to \qcs{L}'$ is a map $\alpha : \gqcs{L} \to \gqcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gqcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gqcs{L'} \dar{\phi'}
  & %\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} 
  & \arrow[swap]{d}{\mu} m^* \gqcs{L} \rar{m^* \alpha} & m^* \gqcs{L'} \dar{\mu'} \\
  \gqcs{L} \rar{\alpha} & \gqcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gqcs{L'} \tight{0}{\boxtimes}{0} \gqcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

The isomorphism $\mu : \bm^* \gqcs{L} \to \LxL$ together with 
\ref{QC.3} guarantee
that the function $\trFrob{\qcs{L}} : G(\Fq)\to \EEx$ is a group homomorphism.  If we write
$\trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$ for pointwise multiplication, then 
 $\trFrob{\qcs{L}\otimes \qcs{L'}} = \trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$.
Since $\qcs{L} \iso \qcs{L'}$ implies $\trFrob{\qcs{L}} = \trFrob{\qcs{L'}}$,
we obtain a group homomorphism
\begin{align*}
\cdef{\trFrob{G}} : \QCiso{G} &\to G(\Fq)^*, \\
\qcs{L} &\mapsto \trFrob{\qcs{L}}.
\end{align*}
This homomorphism is functorial in the following sense:

@@

\begin{proof}
  Let $\qcs{L}$ be a quasicharacter sheaf on $G$. We start by showing that
  $\bf^*\gqcs{L}$ is a local system of rank~one. Let $c_j :\bH^j \hookrightarrow \bH$ be any
  connected component and let $d_i : \bG^i \hookrightarrow \bG$ be the image of $j$ under the group
  homomorphism $\pi_0(\bf) : \pi_0(\bH) \to \pi_0(\bG)$.  Write $\bf^j : \bH^j \to \bG^i$
  for the restriction of $f$ to $\bH^j$.  Since $\gqcs{L}$ is locally constant on the connected
  components of $\bG$, there is a finite \'etale covering
  $a_i : X_i \to \bG^i$ such that $a_i^* (\gqcs{L}\vert_{\bG^i})$
  is constant.  Define $Y_j$ in the following diagram through pullback:
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   Y_j \dar[dashed]{g_j} \rar[dashed]{b_j} & \bH^j \dar{\bf^j} \rar[hook]{c_j} & \bH \dar{\bf} \\
   X_i \rar{a_i} & \bG^i \rar[hook]{d_i} & \bG
  \end{tikzcd}
  \]
  Then $b_j$ is a finite \'etale covering
  of $\bH^j$ and
  \[
  b_j^* \left( (\bf^* \gqcs{L})\vert_{\bH^j}\right)
  = (\bf\circ c_j \circ b_j)^*\gqcs{L}
  = (d_i\circ a_i\circ g_j)^*\gqcs{L} = g_j^* a_i^* (\gqcs{L}\vert_{\bG^i}).
  \]
%
  Since $a_i^* (\gqcs{L}\vert_{\bG^i})$ is a constant sheaf, its pullback
  $b_j^* \left( (\bf^*\gqcs{L})\vert_{\bH^j}\right)$ along $g_j$ is also constant.
  Thus $\bf^*\gqcs{L}$ is locally constant on
  $\bH^j$. To complete the proof that $\bf^*\gqcs{L}$ satisfies 
  \ref{qc.4}
  simply observe that $\bf^*\gqcs{L}$ has rank~one since 
  $(\bf^*\gqcs{L})_{\bg} = \gqcs{L}_{f(\bg)}$, for every geometric point ${\bg}$ on $\bG$.

  To see that $(\bfxf)^* \mu$ satisfies \ref{QC.2}, 
  apply the functor $(\bfxf)^*$
  to \ref{QC.2} for $\qcs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gqcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gqcs{L}$.
  To show that $f^*\qcs{L}$ satisfies 
  \ref{QC.3}, apply the same functor to \ref{QC.3} for $\qcs{L}$.
  The fact that $f$ is a morphism of group schemes defined over $\Fq$ then
  yields isomorphisms $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG}^* (\bfxf)^*$
  and $(\bfxf)^* \bm^*\iso \bm^* \bf^*$ between functors of constructible sheaves.

  Applying $\bf^*$ and $\bf^* \tight{1}{\times}{1}\bf^*$ to \ref{QC.4} defines the action
  of $f^*$ on morphisms of quasicharacter sheaves; arguing as above shows that $f^*$ is
  a functor from $\QC(G)$ to $\QC(H)$.  Since tensor products commute with pullback in schemes,
  $f^* : \QC(G) \to \QC(H)$ is a monoidal functor.
%
  The diagram relating $f^* : \QC(G) \to \QC(H)$, $f^* : G(k)^* \to H(k)^*$ and trace of Frobenius
  commutes by \cite{laumon:87a}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.

@@

@@

Now, let $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$. 
Set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$;
by Proposition~\ref{prop:pullback}, $\qcs{L}_1$ a quasicharacter sheaf on $G_1$ 
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
%
Applying the functor $f^*$ to the isomorphism $\mu$ yields
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already observed that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$.
On the other hand, $f^*(\gqcs{L}\boxtimes \gqcs{L}) = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L}$
by definition.  Since $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L}$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L}$
we get that $f^*(\gqcs{L}\boxtimes \gqcs{L}) = \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies 
\ref{qc.3} as it applies here,
from which it follows that we have exhibited an isomorphism 
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

If $\qcs{L} \ceq (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$, we may define
a quasicharacter sheaf $\qcs{L}' \ceq (\gqcs{L}, \mu, \phi_{k'})$ on the base change
$G_{k'}$ of $G$ to $k'$ by setting
\[
\phi_{k'} \ceq \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (QC.3) for $\phi_{k'}$ 
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gqcs{L}$ 
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
 The rule $\nu_{k'/k}: (\gqcs{L}, \mu, \phi) \mapsto (\gqcs{L}, \mu, \phi_{k'})$ 
 defines a monoidal functor $\QC(G) \to \QC(G_{k'})$. 
 Moreover, the following diagram commutes:
\[
\begin{tikzcd}[column sep=60]
\QCiso{G} \rar{\nu_{k'/k}} \dar{\trFrob{G}} & \QCiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
G(k)^*  \rar{\Nm^*} & G(k')^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
Let $\qcs{L} \ceq (\gqcs{L}, \mu, \phi) \in \QC(G)$ and write $F$ for $\Frob{G}$.  For any $x \in G(k')$,
we may compute the value of $t_{G_{k'}}(\nu_{k'/k}\qcs{L})(x)= t_{\nu_{k'/k}\qcs{L}}(x)$ as the trace of $\phi_{k'}$ on $\gqcs{L}_x$,
and the value of $\Nm^*(t_G(\qcs{L}))(x)$ as the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.
Applying \ref{QC.3} to the stalk of $\gqcs{L}^{\boxtimes n}$ at the point $(x, \Frob{}(x), \ldots, \Frob{}^{n-1}(x))$ yields a diagram
\[
\begin{tikzcd}
\gqcs{L}_{\Nm(x)} \rar \dar{\phi_{\Nm(x)}} & \gqcs{L}_{F(x)} \otimes \gqcs{L}_{F^2(x)} \otimes \cdots \otimes \gqcs{L}_x
\dar{\phi_x \otimes (F^*\phi)_x \otimes \cdots \otimes ((F^{n-1})^*\phi)_x} \\
\gqcs{L}_{\Nm(x)} \rar & \gqcs{L}_x \otimes \gqcs{L}_{F(x)} \otimes \cdots \otimes \gqcs{L}_{F^{n-1}(x)}.
\end{tikzcd}
\]
Choose a basis vector $v$ for $\gqcs{L}_{\Nm(x)}$ and write the image of $v$ under the
bottom map as $v_0 \otimes v_1 \otimes \cdots \otimes v_{n-1}$
for $v_i \in \gqcs{L}_{\Frob{}^i(x)}$.  By \ref{QC.2}, $v$ maps to
$v_1 \otimes v_2 \otimes \cdots \otimes v_0$ along the top of the diagram.
Let $\alpha_i \in \EEx$ represent $((F^i)^*\phi)_x$ with respect to these bases and let $\alpha$ be
the trace of $\phi_{\Nm(x)}$.  We may now equate the trace $\alpha$ of $\phi$ on $\gqcs{L}_{\Nm(x)}$
with the product $\alpha_0 \cdots \alpha_{n-1}$, which is the trace of $\phi_{k'}$ on $\gqcs{L}_x$.
\end{proof}

@@

\begin{definition}
Let $\QCb(G)$ be the category of pairs $(\qcs{L}_0,\mu_0)$ 
where $\qcs{L}_0$ an $\ell$-adic local system on $G$ of rank~one, 
equipped with an isomorphism $\mu_0 : m^* \qcs{L}_0 \to \qcs{L}_0 \boxtimes \qcs{L}_0$ 
satisfying the analogue of \ref{QC.1} on $G$; 
morphisms in $\QCb(G)$ are defined as in the second part of 
\ref{QC.3}. 
This is the category of \emph{bounded quasicharacter sheaves} on $G$. 
\end{definition}

>>>>

We close the introduction with a more detailed description of the structure of this paper.
In Section~\ref{sec:category} we define quasicharacter sheaves and give an interpretation of
them as Weil sheaves with certain properties.
Section~\ref{sec:Frob} introduces the trace of Frobenius homomorphism,
one direction of the sheaf--function dictionary.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on the trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded and finite quasicharacter sheaves,
two subcategories of $\QC(G)$.  These alternate descriptions of quasicharacter sheaves play a key
role in the proof of Proposition~\ref{prop:restriction}.
In section \ref{sec:connected} we show that bounded and finite quasicharacter sheaves are the same
as standard quasicharacter sheaves when $G$ is connected, and that trace of Frobenius is an isomorphism.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and give a more concrete interpretation
of quasicharacter sheaves in this case.  Using a notion of global sections for quasicharacter sheaves,
we give isomorphisms between $\QCiso{G}$, $\Hh^1(\Weil{\Fq}, \Hom(G(\bFq), \EEx))$ and $\Hom(G(\Fq), \EEx)$.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
Section~\ref{sec:snake} closes out Part \ref{part1} with the proof of the main theorem: trace of Frobenius
is an isomorphism for $G$ as long as the geometric component group is finitely generated.

@@

Let $\Frob{}$ denote the geometric Frobenius element in $\Gal(\bFq/\Fq)$ as
well as the corresponding automorphism of $\Spec{\bFq}$. The Weil group
$\Weil{\Fq}\subset \Gal(\bFq/\Fq)$ is the subgroup generated by $\Frob{}$.
Let $\Frob{G} \ceq \id_{G} \times \Frob{}$ be the Frobenius automorphism of $\bG$.
When no other fields play a role we will write $\Frob{}$ and $\Weil{}$ instead.

@@

\begin{definition}\label{def:QC}
A \emph{quasicharacter sheaf on $G$} is a triple
$\qcs{L}\ceq (\gqcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.1)}{CS.1} $\gqcs{L}$ is a rank~one $\ell$-adic local system on $\bG$, by which we mean
a constructible $\ell$-adic sheaf on $\bG$, locally constant on each connected component of $\bG$, whose stalks are one-dimensional $\EE$-vector spaces; 
\labitem{(CS.2)}{CS.2} $\mu: \bm^* \gqcs{L} \to \LxL$ is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gqcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gqcs{L} \boxtimes \gqcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gqcs{L} \boxtimes \bm^* \gqcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gqcs{L} \tight{0}{\boxtimes}{0} \LxL
  \end{tikzcd}
  \]
%
\labitem{(CS.3)}{CS.3} $\phi : \Frob{G}^* \gqcs{L} \to \gqcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gqcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gqcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gqcs{L}\boxtimes \Frob{G}^*\gqcs{L} \dar{\phi\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gqcs{L} \arrow{rr}{\mu}
    && \LxL
  \end{tikzcd}
  \]
\end{enumerate}
\end{definition}

Morphisms in the category of quasicharacter sheaves on $G$, denoted by $\QC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(CS.4)}{CS.4} if $\qcs{L} = (\gqcs{L},\mu,\phi)$ and
  $\qcs{L'} = (\gqcs{L'},\mu',\phi')$ are quasicharacter sheaves on $G$ then
  a morphism $\qcs{L} \to \qcs{L}'$ is a map $\alpha : \gqcs{L} \to \gqcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gqcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gqcs{L'} \dar{\phi'}
  & %\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} 
  & \arrow[swap]{d}{\mu} m^* \gqcs{L} \rar{m^* \alpha} & m^* \gqcs{L'} \dar{\mu'} \\
  \gqcs{L} \rar{\alpha} & \gqcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gqcs{L'} \tight{0}{\boxtimes}{0} \gqcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

The isomorphism $\mu : \bm^* \gqcs{L} \to \LxL$ together with 
\ref{CS.3} guarantee
that the function $\trFrob{\qcs{L}} : G(\Fq)\to \EEx$ is a group homomorphism.  If we write
$\trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$ for pointwise multiplication, then 
 $\trFrob{\qcs{L}\otimes \qcs{L'}} = \trFrob{\qcs{L}} \cdot \trFrob{\qcs{L'}}$.
Since $\qcs{L} \iso \qcs{L'}$ implies $\trFrob{\qcs{L}} = \trFrob{\qcs{L'}}$,
we obtain a group homomorphism
\begin{align*}
\cdef{\trFrob{G}} : \QCiso{G} &\to G(\Fq)^*, \\
\qcs{L} &\mapsto \trFrob{\qcs{L}}.
\end{align*}
This homomorphism is functorial in the following sense:

@@

\begin{proof}
  Let $\qcs{L}$ be a quasicharacter sheaf on $G$. We start by showing that
  $\bf^*\gqcs{L}$ is a local system of rank~one. Let $c_j :\bH^j \hookrightarrow \bH$ be any
  connected component and let $d_i : \bG^i \hookrightarrow \bG$ be the image of $j$ under the group
  homomorphism $\pi_0(\bf) : \pi_0(\bH) \to \pi_0(\bG)$.  Write $\bf^j : \bH^j \to \bG^i$
  for the restriction of $f$ to $\bH^j$.  Since $\gqcs{L}$ is locally constant on the connected
  components of $\bG$, there is a finite \'etale covering
  $a_i : X_i \to \bG^i$ such that $a_i^* (\gqcs{L}\vert_{\bG^i})$
  is constant.  Define $Y_j$ in the following diagram through pullback:
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   Y_j \dar[dashed]{g_j} \rar[dashed]{b_j} & \bH^j \dar{\bf^j} \rar[hook]{c_j} & \bH \dar{\bf} \\
   X_i \rar{a_i} & \bG^i \rar[hook]{d_i} & \bG
  \end{tikzcd}
  \]
  Then $b_j$ is a finite \'etale covering
  of $\bH^j$ and
  \[
  b_j^* \left( (\bf^* \gqcs{L})\vert_{\bH^j}\right)
  = (\bf\circ c_j \circ b_j)^*\gqcs{L}
  = (d_i\circ a_i\circ g_j)^*\gqcs{L} = g_j^* a_i^* (\gqcs{L}\vert_{\bG^i}).
  \]
%
  Since $a_i^* (\gqcs{L}\vert_{\bG^i})$ is a constant sheaf, its pullback
  $b_j^* \left( (\bf^*\gqcs{L})\vert_{\bH^j}\right)$ along $g_j$ is also constant.
  Thus $\bf^*\gqcs{L}$ is locally constant on
  $\bH^j$. To complete the proof that $\bf^*\gqcs{L}$ satisfies 
  \ref{cs.4}
  simply observe that $\bf^*\gqcs{L}$ has rank~one since 
  $(\bf^*\gqcs{L})_{\bg} = \gqcs{L}_{f(\bg)}$, for every geometric point ${\bg}$ on $\bG$.

  To see that $(\bfxf)^* \mu$ satisfies \ref{CS.2}, 
  apply the functor $(\bfxf)^*$
  to \ref{CS.2} for $\qcs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gqcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gqcs{L}$.
  To show that $f^*\qcs{L}$ satisfies 
  \ref{CS.3}, apply the same functor to \ref{CS.3} for $\qcs{L}$.
  The fact that $f$ is a morphism of group schemes defined over $\Fq$ then
  yields isomorphisms $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG}^* (\bfxf)^*$
  and $(\bfxf)^* \bm^*\iso \bm^* \bf^*$ between functors of constructible sheaves.

  Applying $\bf^*$ and $\bf^* \tight{1}{\times}{1}\bf^*$ to \ref{CS.4} defines the action
  of $f^*$ on morphisms of quasicharacter sheaves; arguing as above shows that $f^*$ is
  a functor from $\QC(G)$ to $\QC(H)$.  Since tensor products commute with pullback in schemes,
  $f^* : \QC(G) \to \QC(H)$ is a monoidal functor.
%
  The diagram relating $f^* : \QC(G) \to \QC(H)$, $f^* : G(k)^* \to H(k)^*$ and trace of Frobenius
  commutes by \cite{laumon:87a}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.

@@

Consider the short exact sequence in the category of $\QC$-schemes
defining the component group scheme for $G$:
\begin{equation}\label{eq:pi0}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is also a $\QC$-scheme, 
Proposition~\ref{prop:pullback} implies that \eqref{eq:pi0} defines a sequence of functors
\[
\begin{tikzcd}
\QC(0) \rar & \QC(\pi_0(G)) \rar & \QC(G) \rar & \QC(G^0) \rar & \QC(0)
\end{tikzcd}
\]
which induces a sequence of group homomorphisms 
\[
\begin{tikzcd}
\QCiso{0} \rar & \QCiso{\pi_0(G)} \rar & \QCiso{G} \rar & \QCiso{G^0} \rar & \QCiso{0}.
\end{tikzcd}
\]
after passing to isomorphism classes.  It is easy to see that $\QCiso{0}$ is trivial. In Sections~\ref{sec:connected} and \ref{sec:etale} we will find the groups $\QCiso{G^0}$
and $\QCiso{\pi_0(G)}$ for all $\QC$-schemes, 
which will allow us to conclude $\QCiso{G} \iso G(\Fq)^*$.

@@

Now, let $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$. 
Set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$;
by Proposition~\ref{prop:pullback}, $\qcs{L}_1$ a quasicharacter sheaf on $G_1$ 
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
%
Applying the functor $f^*$ to the isomorphism $\mu$ yields
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already observed that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$.
On the other hand, $f^*(\gqcs{L}\boxtimes \gqcs{L}) = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L}$
by definition.  Since $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L}$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L}$
we get that $f^*(\gqcs{L}\boxtimes \gqcs{L}) = \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies 
\ref{CS.3} as it applies here,
from which it follows that we have exhibited an isomorphism 
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

If $\qcs{L} \ceq (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$, we may define
a quasicharacter sheaf $\qcs{L}' \ceq (\gqcs{L}, \mu, \phi_{k'})$ on the base change
$G_{k'}$ of $G$ to $k'$ by setting
\[
\phi_{k'} \ceq \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (CS.3) for $\phi_{k'}$ 
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gqcs{L}$ 
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
 The rule $\nu_{k'/k}: (\gqcs{L}, \mu, \phi) \mapsto (\gqcs{L}, \mu, \phi_{k'})$ 
 defines a monoidal functor $\QC(G) \to \QC(G_{k'})$. 
 Moreover, the following diagram commutes:
\[
\begin{tikzcd}[column sep=60]
\QCiso{G} \rar{\nu_{k'/k}} \dar{\trFrob{G}} & \QCiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
G(k)^*  \rar{\Nm^*} & G(k')^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
Let $\qcs{L} \ceq (\gqcs{L}, \mu, \phi) \in \QC(G)$ and write $F$ for $\Frob{G}$.  For any $x \in G(k')$,
we may compute the value of $t_{G_{k'}}(\nu_{k'/k}\qcs{L})(x)= t_{\nu_{k'/k}\qcs{L}}(x)$ as the trace of $\phi_{k'}$ on $\gqcs{L}_x$,
and the value of $\Nm^*(t_G(\qcs{L}))(x)$ as the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.
Applying \ref{CS.3} to the stalk of $\gqcs{L}^{\boxtimes n}$ at the point $(x, \Frob{}(x), \ldots, \Frob{}^{n-1}(x))$ yields a diagram
\[
\begin{tikzcd}
\gqcs{L}_{\Nm(x)} \rar \dar{\phi_{\Nm(x)}} & \gqcs{L}_{F(x)} \otimes \gqcs{L}_{F^2(x)} \otimes \cdots \otimes \gqcs{L}_x
\dar{\phi_x \otimes (F^*\phi)_x \otimes \cdots \otimes ((F^{n-1})^*\phi)_x} \\
\gqcs{L}_{\Nm(x)} \rar & \gqcs{L}_x \otimes \gqcs{L}_{F(x)} \otimes \cdots \otimes \gqcs{L}_{F^{n-1}(x)}.
\end{tikzcd}
\]
Choose a basis vector $v$ for $\gqcs{L}_{\Nm(x)}$ and write the image of $v$ under the
bottom map as $v_0 \otimes v_1 \otimes \cdots \otimes v_{n-1}$
for $v_i \in \gqcs{L}_{\Frob{}^i(x)}$.  By \ref{CS.2}, $v$ maps to
$v_1 \otimes v_2 \otimes \cdots \otimes v_0$ along the top of the diagram.
Let $\alpha_i \in \EEx$ represent $((F^i)^*\phi)_x$ with respect to these bases and let $\alpha$ be
the trace of $\phi_{\Nm(x)}$.  We may now equate the trace $\alpha$ of $\phi$ on $\gqcs{L}_{\Nm(x)}$
with the product $\alpha_0 \cdots \alpha_{n-1}$, which is the trace of $\phi_{k'}$ on $\gqcs{L}_x$.
\end{proof}

@@

\begin{definition}
Let $\QCb(G)$ be the category of pairs $(\qcs{L}_0,\mu_0)$ 
where $\qcs{L}_0$ an $\ell$-adic local system on $G$ of rank~one, 
equipped with an isomorphism $\mu_0 : m^* \qcs{L}_0 \to \qcs{L}_0 \boxtimes \qcs{L}_0$ 
satisfying the analogue of \ref{CS.1} on $G$; 
morphisms in $\QCb(G)$ are defined as in the second part of 
\ref{CS.3}. 
This is the category of \emph{bounded quasicharacter sheaves} on $G$. 
\end{definition}
